---
title: "TS - Varicella Forecast"
author: "Benjamin BERNARD"
date: "28/10/2021"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First we start to load necessary packages and dateset

```{r}
library(forecast)
library(ggplot2)
data=read.csv(file="http://eric.univ-lyon2.fr/~jjacques/Download/DataSet/varicelle.csv")
varicelle<-ts(data$x,start=c(1931,1),end=c(1972,6),freq=12)
plot(varicelle)

```

1\. **Data Manipulation**

This is the first step to understand the dataset we have.

When looking to the plot, data looks seasonal. Seasonal means that there is a correlation between the time and the value to analyse. In other words, same even happen at the same period.

To visualise that we can plot a correlogram :

```{r}
tmp=acf(varicelle, type="cor", plot=TRUE)
ptmp=pacf(varicelle, type="cor", plot=FALSE)
plot (ptmp)

```

Correlogram shows the correlations. There are significant when it goes above or below the blue dotted line. As the correlations are periodic here (looks that the ACF is sinusoidal) we can conclude that the data are correlated on a cyclic pattern and so are seasonal.

```{r}
ggseasonplot(varicelle,year.labels= TRUE,year.labels.left=TRUE)
```

The seasonal plot above confirm the seasonality : we can easily see a peak February and April and a dramatic decreasing with a minimum value in September.

By computing the values year per year we can see better that on top of the seasonality we have a decreasing trend : varicella number of cases is decreasing over the time.

```{r}
x=rep(0,41)
for (i in 0:40) x[i+1]<-sum(varicelle[(1+12*i):(12*(i+1))])
plot(x,type='l',xaxt='n',xlab='')
axis(1,at = 0:40,labels = 1931:1971)
```

2.  **Forecasting**

As usual we split the time series in test and train subsets :

```{r}
serie_train=window(varicelle,start=c(1931,1),end=c(1970,12))
serie_test=window(varicelle,start=c(1971,1),end=c(1972,6))

```

How to choose the size of the test set ? It should be the same size than the period you want to forecast : if you want to forecast 10 years, then your test set must be 10 years long.

Now we'll try several forecast models and find the best one :

**Holt Winters exponential smoothing Model**

```{r}
fithw=hw(serie_train,lambda="auto")
prevhw=forecast(fithw,h=18)
autoplot(prevhw, xlim=c(1965,1972.6),ylim=c(0,2500)) + autolayer(serie_test, series="true data")+
autolayer(prevhw$mean, series="HW forecasts")
```

**Automatic SARIMA Model**

The parameters of this model are defined automatically by the function.

```{r}
fitasm=auto.arima(serie_train,lambda="auto")
prevasm=forecast(fitasm,h=18)
autoplot(prevasm,xlim=c(1965,1972.6),ylim=c(0,2500)) + autolayer(serie_test, series="true data")+
autolayer(prevasm$mean, series="Auto SARIMA forecasts")


```

System choose an ARIMA(1,0,0)(1,1,1)[12] model. We'll compare it later with the other model tried.

**Auto-regressive Neural Network**

Now we'll use a deep learning model.

```{r}
fitnn=nnetar(serie_train,lambda = "auto")
prevnn=forecast(fitnn,h=18)
autoplot(prevnn,xlim=c(1965,1972.6),ylim=c(0,2500)) + autolayer(serie_test, series="true data")+
autolayer(prevnn$mean, series="Neural Network Forecast")

```

**Forecasting models comparison**

```{r}
cat('RMSE Hot Winters :',sqrt(mean((serie_test-prevhw$mean)^2)),'\n')
cat('RMSE Automatic SARIMA:',sqrt(mean((serie_test-prevasm$mean)^2)),'\n')
cat('RMSE Neural Network:',sqrt(mean((serie_test-prevnn$mean)^2)),'\n')
```

We have a winner : Automatic SARIMA Model ! But this can be improved.

3.  **Model improvement**
