---
title: "TS - San Francisco Precipitations"
author: "Benjamin BERNARD"
date: "28/10/2021"
output: rmarkdown::github_document
---

On this page we'll forecast precipitations for the city of San Francisco.

1.  **Dataset splitting**

Let's download and split the dataset into train and test sets :

```{r}
data=scan(file="http://eric.univ-lyon2.fr/~jjacques/Download/DataSet/sanfran.dat",skip=1)
sanfran<-ts(data,start=c(1932,1),end=c(1966,12),freq=12)
library(forecast)
sanfran_train=window(sanfran,,start=c(1932,1),end=c(1963,12))
sanfran_test=window(sanfran,,start=c(1964,1),end=c(1966,12))
```

2.  **Data study and Model configuration**

Now we can visualize both subsets

```{r}
plot(sanfran_train,xlim=c(1932,1970),ylim=c(0,120))
lines(sanfran_test,lty=2)
```

Looks like we have a seasonal pattern here.

Let's create a model and try it on the test set :

```{r}
library(forecast)
h=hw(sanfran_train,damped=FALSE,h=36)
plot(sanfran_train,xlim=c(1964,1967),ylim=c(0,120))
lines(sanfran_test,lty=2)
lines(h$mean,col=2)
print(sqrt(mean((h$mean-sanfran_test)^2)))
accuracy(h,sanfran_test)
h
```

With this model the RMSE (Root mean square error) is : 15.99891.

Obviously, we can also see if the model fits well or not just comparing the plots.

Let's try a damped version to see if it fits better our test set.

```{r}
hd=hw(sanfran_train,damped=TRUE,h=36)
plot(sanfran_train,xlim=c(1964,1967),ylim=c(0,120))
lines(sanfran_test,lty=2)
lines(hd$mean,col=3)
print(sqrt(mean((hd$mean-sanfran_test)^2)))
hd
```

We can see that the RMSE is slightly better. So we'll use this damped version to forecast unknown precipitations values.

Let's see that adding both models to the same plot :

```{r}
plot(sanfran_train,xlim=c(1966,1967),ylim=c(10,70))
lines(sanfran_test,lty=2)
lines(hd$mean,col=3)
lines(h$mean,col=4)
```

Not easy to see any difference between damped and undamped version. Hopefully we may compute the RMSE to compare both and find the best model.

3.  **Forecast**

Now we'll try to forecast the values after 1967, where our dataset stops.

```{r}
hf=hw(sanfran,seasonal='additive',damped=TRUE, h=24)
plot (sanfran,xlim=c(1966,1969),ylim=c(0,120))+lines(hf$mean, col=5)

```
