---
title: "TS - O2 Forecast"
author: "Benjamin BERNARD"
date: "28/10/2021"
output: rmarkdown::github_document
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1.  **Dataset splitting**

The test subset must have the size of the data we want to forecast.
Here we want to predict the next 7 years so we keep only the last 7 years for the train subset.

```{r}
co2_train=window(co2,start=c(1959,1),end=c(1989,12))
co2_test=window(co2,start=c(1990,1),end=c(1997,12))
```

We can plot both subset.

```{r}
plot(co2_train,xlim=c(1960,2010),ylim=c(300,400))
lines(co2_test,lty=2)
```

2.  **Visual Analyse**

We can see on the above plot that we have the following characteristics :

-   A seasonality (values looks to have the same behavior over the time - increasing and decreasing)

-   A trend (values grows along the dataset)

3.  **Forecasting**

Data looks quite easy to forecast.
Let's try with the HoltWinters forecasting method.

To create the model, we use hx() function with the following parameters :

-   dataset to use (here for this step we will use co2_train as it's our training model)

-   seasonal='additive' as the trend is increasing proportionally.
    We could use 'multiplicative' in case of an exponential trend.

-   h is the number of values to forecast after the dataset we choose.
    96 here is the amount of values in the test set.

Then we'll plot the whole dataset (co2_train in plain line, and co2_test in dotted line) plus the forecast to visualise the quality of our forecast.

```{r}
#install.packages("forecast")
library(forecast)
h=hw(co2_train,seasonal='additive',h=96)
plot(co2_train,xlim=c(1985,2002),ylim=c(340,380))
lines(co2_test,lty=2)
lines(h$mean,col=2)
accuracy(h)
```

We may try to use a Damped version :

```{r}
hd=hw(co2_train,seasonal='additive',damped=TRUE,h=96)
plot(co2_train,xlim=c(1985,2002),ylim=c(340,380))
lines(co2_test,lty=2)
lines(hd$mean,col=2)
lines(hd$mean,col=3)
print(sqrt(mean((hd$mean-co2_test)^2)))
```

We can see that it's worse... So we will use our first model which forecast reasonably our dataset.
Let's forecast 10 years.

```{r}
h=hw(co2,seasonal='additive',damped=FALSE,h=120)
autoplot(co2,xlim=c(1985,2011),ylim=c(320,400))+autolayer(h)
h
```

Great job, we predicted unknown values in the future of our dataset !
